---
title: "pacdev_assignment_PM"
author: "Prabhudatta Mohapatra"
date: "03/26/2023"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
#loading libraries
library(tidyverse)
library(ggplot2)
library(dplyr)
```

```{r}
#importing data
pacd<- read.csv("pacdev_data.csv")
#checking data summary
summary(pacd)
str(pacd)
```


## Q1

```{r}
#liner regression summary
summary(lm(price~sqft,data = pacd))
```
```{r}
#price and sqft plot with summary linear regression line
ggplot(pacd,aes(sqft,price)) +
  geom_jitter() +
  geom_smooth(formula = y~ x,method = 'lm',se = F) +
  labs(title = "Plot between price and sqft")
```


> From the above summary of the linear regression we can write the equation of the summary line as: price = (269.345 * sqft) + 40623.019. As the sqft co-efficient is positive it signifies that there is a positive relatinoship between sqft and price or as sqft increases price of home increases. Same can be derived looking at the regressino line positive slope. Also, sqft co-efficient 269.345 suggests that 1 unit increase in sqft is associated with an increase of $269.35 in price of the home, on average. The coefficient for sqft (269.345) is statistically significant at the p < 0.001 level. For intercept the coefficient is 40623.019 is the average value of price when there sqft is zero (0), which also states that the home prices strats at $40623.02, on average. 

## Q2

```{r}
#centering the sqft
pacd<- pacd %>% 
  mutate(centered_sqft = (sqft - mean(sqft))) 

#multiple regression model
summary(lm(price ~ city + pool + garage  + bed + bath + centered_sqft, data = pacd))
```

> From the above model summary the equation of the regression line: price = 528103.213 + (190239.704 * citySanta Monica) + (88020.719 * cityWestwood) + (10124.630 * poolyes) - (14195.911 * garagyes) + (41.553 * bed) - (3092.909 * bath) + (271.561 * centered_sqft)

>Interpretation of coefficients:

> `Intercept`: 528103.213 is average price for homes when city = Long Beach, Pool = No, garage = No, bed = 0, bath = 0, and centered_sqft = 0 (which states average the starting price of homes considering adjusted variables) 

> `centered_sqft`:   an increase of 1 sqft is associated with an increase of $271.56 in the price of home, on average 

> `citySanta Monica`: 190239.704 is the average difference in home price comparing Long Beach to Santa Monica  

> `cityWestwood`: 88020.719 is the average difference in home price comparing Long Beach to Westwood. 


## Q3

```{r}
#interaction model
summary(lm(price ~ pool + garage  + bed + bath + centered_sqft * city, data = pacd))
```

```{r}
ggplot(pacd,aes(centered_sqft, price, col = city)) + 
  geom_jitter() + 
  geom_smooth(formula= y ~ x, method = 'lm', se = FALSE) + 
  labs(title = "Plot of the interaction model (price ~ centered_sqft * city)")
```

> Above summary and graph represents home sale price model having an interaction between centered_sqft and city.  

>Interpretation of coefficients:

> `Intercept`: 527445.45 is average price for homes when city = Long Beach, Pool = No, garage = No, bed = 0, bath = 0, and centered_sqft = 0 (which states average starting price of homes considering adjusted variables) 

>  `centered_sqft`:   as per the interaction model an increase of 1 sqft is associated with an increase of $240.63 in the price of home, on average. It is statistically significant. 

>  `citySanta Monica`: 189904.29 is the average difference in home price comparing Long Beach to Santa Monica. It is statistically significant.  

> `cityWestwood`: 88037.11 is the average difference in home price comparing Long Beach to Westwood. It is statistically significant.

> `centered_sqft:citySanta Monica`: 90.09 is the average difference in the relationship between centered_sqft and home sale price for Long Beach vs. Santa Monica. The relationship can be stated the other way around also:  90.09 is the average difference in the relationship between city (Long Beach vs. Santa Monica) and home sale price for 1 unit change in the centered_sqft. The coefficient 90.09 estimate the change in the slope of the regression line for Santa Mnica compared to Long Beach. As the coefficient is positive that means the regression line relating sqft to price is steeper for Santa Monica in comparison to Long Beach and there is a strong relationship. . It is statistically significant.

> `centered_sqft:cityWestwood`: 37.95 is the average difference in the relationship between centered_sqft and home sale price for Long Beach vs. Westwood. The relationship can be stated the other way around also:  37.95 is the average difference in the relationship between city (Westwood vs. Santa Monica) and home sale price for 1 unit change in the centered_sqft. The coefficient 37.95 estimate the change in the slope of the regression line for Westwood compared to Long Beach. As the coefficient is positive that means the regression line relating sqft to price is steeper for Westwood in comparison to Long Beach and there is a strong relationship.

> Also, it can be observed that Santa Monica city interaction with centered_sqft has the highest interaction effect for price when compared to other cities (Long Beach, Westwood) combination with centered_sqft. Same can be observed from the interaction model plot that Santa Monica city home sale price has the highest positive slope (price and centered_sqft*city) among all the cities. 

## Q4

```{r}
# Model assessment
#residual plot
pacd %>% 
  mutate(fitted = fitted(lm(price ~ pool + garage  + bed + bath + centered_sqft * city, data = pacd)),
      residuals = fitted-price) %>% 
  ggplot(aes(fitted, residuals)) +
  geom_point() +
  geom_smooth(formula=y~x,method = "lm", se = F, col = "blue") +
  labs(title = "Residuals for the interaction model")
```

>From the residual plot we can observe that the residuals are randomly distributed around the summary line  or residual line = 0. Also, from visual inspection we do not see any pattern in residual distribution which satisfies the condition for selecting the correct regression technique for modeling. So, we can conclude that a linear regression model is appropriate for modeling this particular data.

```{r}
#extracting fitted values
fitted_values<-fitted(lm(price ~ pool + garage  + bed + bath + centered_sqft * city, data = pacd))  

#fitted values vs observed values plot for price
ggplot(pacd,aes(fitted_values, price)) +
  geom_point() +
  geom_smooth(formula=y~x,method = "lm", se = F, col = "blue")+
  labs(title = "Plot between fitted and observed price values")
```

>From the above plot we can observe the linearity assumption condition of the data is satisfied and so, implemented linear regression method is justified. Also, as the data is linear, data is accurately presented with a line that satisfies the conditions of a good model. We can see that most of the points are pretty close to the summary line, however some of the points are away from summary line, which tells that may be we can transform data to get a better fit. 


```{r}
residual<-resid(lm(price ~ pool + garage  + bed + bath + centered_sqft * city, data = pacd))
# histogram of residuals
ggplot(pacd, aes(residual)) +
  geom_histogram(bins = 30) +
  labs(title = "Histogram of residuals")
```

>Above plot shows the histogram of the residuals. Residual histogram is used to verify normal distribtution of the variance. A close observation of the plot shows a symmetric bell-shaped histogram that is evenly distributed around zero indicates that the residuala are normally distributed and the assumption that variance is normally distributted is true. This also confirms that the implemented linear regression model is the right model for the data.

## Q5

>Considering the PacDev's business model of remodelling and reselling homes with a primary focus on increasing square footage, Andrew should recommend that PacDev prioritize remodelling and reselling homes in the city of Santa Monica over those in Long Beach and Westwood becuase in Santa Monica adding square footage has a strong effect on home price increase. From the interaction model we can observe that interaction between centered_sqft:citySanta Monica has an effect size of 90.09, which is positive and signifies a stronger relationship. Though the interaction between centered_sqft:cityWestwood is positive (37.95), it is almost 3 times less than that of centered_sqft:citySanta Monica. Same conclusion can be derived from the interaction model plot that Santa Monica has the highest positive slope among all the cities. Comparing different regression models, Andrew shoud recommend the interaction model (if PacDev wants to predict the price or want to add more data or variables to rerun the analysis) as the interaction mode has the best fit among all the models designed having the highest R-square value of 0.31 and the lowest Residual standard error of 194500. To implement a statistical model in real world it should satisfy standard terms of a good model which were checked and confirmed such as model should have linearlity of the relationship between outcome and predictor variables (confirmed through observed vs fitted values plot), independence of the errors (confirmed through residuls vs fitted values plot: no pattern observed), normality of errors and equal variance of errors (confirmed through residuals histogram: normal distribution of residuals), and low residual standard error and high R-square in the interaction model comparing other models. Hence, Andrew should recommend pacDev to prioritize business in Santa Monica to maximize price increase for additional square footage as Santa Monica will have the highest increase in home sale price by adding square footage in comparison to other cities       


## Challenge

```{r}

```

