---
title: "Velo.com_Assignment_PM"
author: "Prabhudatta Mohapatra"
date: "03/12/2023"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

<!-- Note:   -->

<!-- These instructions are commented out and will not display when you knit your RMarkdown document. -->

<!-- - Change the information in the yaml header above:  title, author, data. -->
<!-- - Make sure output is html_document. -->
<!-- - Once you are finished coding, **run each chunk individually to make sure there are no errors**.  (If necessary fix your code.) Once your code is error-free, click "knit" on the menu above. Your document should compile to HTML, provided that you have output set to "html_document." -->
<!-- - In the code chunk above ("setup") echo is set to TRUE.  This means that the code in your chunks will be displayed, along with the results, in your compiled document. -->

## Load data and packages

```{r}
#loading packages
library(tidyverse)
library(dplyr)
library(ggplot2)
v <- read_csv("velo.csv")
summary(v)
```

## Questions

Lightly comment your code and use pipes for readability.

Comment briefly on each of the questions, as directed.  Only the the final question requires a lengthier response.

### Q1

Plot the distribution of `spent` by `checkout_system`.  Below you will use a t-test to compare these distributions statistically.  However, a t-test assumes normally distributed data.  Is that assumption valid in this case?  Why or why not?

Note: 

1. You could compare the two distributions using histograms but a density plot works better. (A boxplot is also an option.)

2. Make sure to include a plot title.

```{r}
#Distribution (Density Plot) of amount spent by check out system
ggplot(v,mapping = aes(x = spent, col = checkout_system)) +
  geom_density() + 
  theme_minimal() +
  labs(title = "The distribution of amoun spent by checkout_system.") + 
  xlim(-5000,10000)
```
```{r}
#Distribution (Histogram) of amount spent by check out system
ggplot(v, aes(x = spent)) +
  geom_histogram() + 
  theme_minimal() +
  facet_wrap(~checkout_system, ncol = 1)  +
  labs(title = "The distribution of amount spent by checkout_system")
```

```{r}
#Distribution (Histogram) of amount spent by check out system
ggplot(v, mapping = (aes(x = checkout_system, y = spent)))+
  geom_boxplot() +
  theme_minimal() 
```

> Answer:  From the above density plot, histogram, and boxplot, we can observe that both distributions appear to be normally distributed as they are bell-shaped. Both distributions look similar and are a little right-skewed. As the t-test uses the mean for group comparison and using the mean implies that the distributions are normally distributed. The assumption is valid for the data amount spent at both checkout_systems because if the distributions are not normally distributed then it isn’t valid to use the p-values from the t-test to compare the differences between the samples. From the below calculation, we can observe that the mean and median are pretty close to each other which indicates that both distributions are normally distributed.

### Q2

Create a summary table of `spent` by `checkout_system` with the following statistics:

- n
- mean
- median
- standard deviation
- total 
- the lower and upper bound of a 95% z-confidence interval for the mean.

Your table should have 2 rows and 8 columns.

```{r}
#Summary table of spent by checkout_system
v %>%
  group_by(checkout_system) %>%
  summarize(n = n(),
            mean = mean(spent),
            median = median(spent),
            sd = sd(spent),
            total = sum(spent),
            lower_bound = (mean - 1.96 * (sd/sqrt(n))) %>% round(2),
            upper_bound = (mean + 1.96 * (sd/sqrt(n))) %>% round(2))
```
> Above summary statistics presents the mean, median, standard deviation, total, lower confidence interval and upper confidence interval of spent for both checkout systems.

###  Q3

Is average spending significantly higher in the treatment group?  (The treatment group consists in the customers using the new checkout system.)  Answer this question using a 2 sample, 2-tailed t-test with alpha set at .05. (Note that these are the default settings for the `t.test()` function when vectors are supplied for the x and y arguments.)

```{r}
# run 2-sample, 2-tailed t-test with standard alpha of .05
t.test(filter(v, checkout_system == "new")$spent,
       filter(v, checkout_system == "old")$spent,
       alternative = c("two.sided"))
```

> Answer: As the t-test provides a p-value of 0.1536, which is greater than 0.05 significance level (p>alpha), we do not reject the null hypothesis (HO: No difference between old and new checkout systems.) and conclude that there is no true difference between the means of spending for checkout systems. The 95 % confidence interval (-23.45 ~ 148.93) includes zero also indicates the same result. Hence, the average spending is not significantly higher in the treatment group.

### Q4

Create another summary table of `spent` by `checkout_system` and `device`.  Include these same statistics:

- n
- mean
- median
- standard deviation
- the lower and upper bound of a 95% confidence interval for the mean.

```{r}
v%>% 
  group_by(device, checkout_system) %>% 
  summarize(n = n(),
            mean = mean(spent),
            median = median(spent),
            sd = sd(spent),
            lowerCI = (mean - 1.96 * (sd/sqrt(n))) %>%  round(2),
            upperCI = (mean + 1.96 * (sd/sqrt(n)))  %>%  round(2))
```

The table should have 4 rows and 8 columns.  

Based on this information (as well as Sarah's observation, noted in the case description, that the glitch in the checkout system seemed more prevalent for mobile users), an additional statistical comparison of new and old among just mobile users seems warranted. Make that comparison using a 2 sample, 2-tailed t-test with alpha set at .05.  Report your results.

Note that a t-test can only compare two groups.  Therefore, you will need to subset the data before making the comparison.

```{r}
new_mobile=subset(v,checkout_system=='new'&device=='mobile')
old_mobile=subset(v,checkout_system=='old'&device=='mobile')
t.test(x =new_mobile$spent,
       y = old_mobile$spent,
       alternative = c("two.sided"))
```

> Answer: The t-test between two groups of mobile users (as per two different checkout systems) provides a p-value of 0.01655 which is less than 0.05 significance level (p<alpha), so we reject the null hypothesis (HO: No difference between old and new checkout systems.) and conclude that there is a true difference between the means of spending for different checkout systems for mobile users. The 95 % confidence interval (27.013 ~ 269.138) does not include zero, which also indicates the same result. Hence the average spending is higher in the treatment group while considering only mobile users.


### Q5

What course of action should Sarah recommend to the management at velo.com? Please incorporate your analytic results from above in fashioning an answer.

> Answer: Based on the above analyses, Sarah should recommend the new checkout system over the old checkout system for mobile users as we found that the average spending is higher in the treatment group (mobile users using the new checkout system) while considering only mobile users. Comparing the amount spent by users of the old vs. new checkout systems on computers and mobile devices, there is no significant difference in average spending (assuming a normal distribution for the amount spent, the t-test returned a p-value of 0.1536, which was greater than significance level threshold of 0.05, not rejecting the null hypothesis). However, when comparing the amount spent by users of the new vs. old systems on mobile devices only, it is found that mobile users of the new system spent ($2322.996) significantly more on average than mobile users of the old system ($2174.920), this result was statistically significant (assuming a normal distribution for the amount spent by the mobile users, the t-test returned a p-value of 0.01655, which was less than our significance level threshold of 0.05, rejecting the null hypothesis). So, as per the A/B test results, comparing spending of mobile users using the new vs. old checkout systems, Sarah should recommend the management of velo.com to adopt the new checkout system for mobile users only as there is a true difference in the means of new and old checkout system samples specifically for mobile users(average spending is higher in the treatment group) and to avoid dropping off of customers during the checkout process of complicated and expensive custom bikes. Adopting a new checkout system for mobile users will improve the overall revenue of velo.com.  
  


### Challenge

In looking at the summary tables you created above you might wonder about differences not just in spending but also in the number of customers.  After all, the case description indicated that customers may have been prevented from completing purchases using the old checkout system. Here are the counts:

```{r}

table(v$checkout_system) 

```

Obviously there are some notable differences in the number of customers  Are these differences statistically significant?

We could answer this question using simulation. For example, the binomial distribution could be used to represent the null distribution, the number of expected buyers under the null hypothesis of no difference between the checkout systems (that is, no difference in buying probability).  The *observed* proportion of buyers under the new system is 1828 / (1828 + 1655) = .525. How often would this proportion occur under the null? 

```{r}
# We will use the rbinom() function to do this simulation. n refers to the number of simulations, 
# size refers to the number of trials, and prob is the probability of getting a 1 under the null. 

# Example:
rbinom(n = 1, size = 1, prob = .5)
rbinom(10, 1, .5)
rbinom(10, 10, .5)

# Here is the simulation.  Note that we divide by the total number of trials to obtain the proportion of 1s.
set.seed(123)
sims <- rbinom(n = 100000, size = nrow(v), prob = .5) / nrow(v)

hist(sims)
```

The observed proportion would not happen very often under the null.  Let's calculate a formal p-value.

```{r}
(sims >= (1828 / (1828 + 1655))) %>% mean
```

We would double this for a 2-sided test, of course, but the result is still easily statistically significant at the conventional threshold of p < .05.   

The Chi-squared test is the statistical test typically used in this situation to do a formal hypothesis test of the counts in a 1 x 2 or 2 x 2 (or larger) contingency table. Here is a Kahn Academy video on it: 

https://www.khanacademy.org/math/ap-statistics/chi-square-tests/chi-square-goodness-fit/v/chi-square-statistic. 

And here is the Wikipedia article:  

https://en.wikipedia.org/wiki/Chi-squared_test. 

Here is the R function:

```{r}

?chisq.test

```

Note that this R function takes a table as its argument:

```{r}

chisq.test(table(v$checkout_system))

```

Notice that the p-value is almost identical to what we calculated using simulation!

1. Explain the chisquare test.
2. Run the chisquare test also on the 2 x 2 contingency table comparing checkout system and device. 
3. Interpret the statistical results for the chisquare tests for both the 1 x 2 table and the 2 x 2 table. 
4. What is the relevance of these for the velo.com case?

